### 2015-05-25 paper/paper.tex
Added Discussion and Conclusion section content

Indexing in MongoDB is supported in a similar ways that are present in relational databases. Mongo builds a B-Tree data
structure on a particular collection's key, allowing all documents within that collection with that key to be accessed
in $O(log n)$ time, as opposed to $O(n)$ time if unindexed. MongoDB also supports numerous types of indexes, such as
single-key, multi-key, geospatial, and hashed indexes.

@@

As most of what has been shown in~\sectref{sec:implementation}, properly setting up the MongoDB database for use with the IRT
project is not a hard task and only requires relatively trivial MongoDB operations to be performed.

@@

>>>>

Indexing in MongoDB is supported in a similar ways that are present in relational databases. Mongo builds a B-Tree data
structure on a particular collection's key, allowing all documents within that collection with that key to be accessed
in $O(log(n))$ time, as opposed to $O(n)$ time if unindexed. MongoDB also supports numerous types of indexes, such as
single-key, multi-key, geospatial, and hashed indexes.

@@

As most of what has been shown in~\sectref{sec:implementation}, properly setting up the MongoDB database for use with the IRT
project is not a hard task and only requires relatively trivial MongoDB operations to be performed. The overall design
decisions for the project is the more difficult part of the implementation, and is highly dependant on the use-case
for the database. Hence, more time generally would be spent on the design of the MongoDB database, and use of the
mongo shell, or appropriate MongoDB driver/API for a specific programming language, can be used to structure everything.

As noted in~\sectref{sec:mongo_overview}, MongoDB allows for much more advanced operations to be performed, however
most of the available functions are easy to learn and offer varying amounts of power over given database collections,
depending how many options the programmer makes use of.

Coming from a object-oriented and functional software engineering background, I found interacting with MongoDB through its provided JavaScript shell
a very easy task, and far more intuitive than using declarative SQL on a traditional relational database. However, this is merely
anecdotal, and for a database user with much experience with SQL and the relational data model, using MongoDB, and
similarly functioning NoSQL databases, may prove to be a difficult task; particularly for those without a lot of imperative programming
experience. However, as a counter-argument, for basic interaction with MongoDB, scripts can easily be written by
following the API documentations without requiring a thorough understanding of the underlying programming languages.

A further difference that may catch many users of relational database, is the lack of joins, as noted in~\sectref{ssub:joins}.
As MongoDB does not natively support multi-document joins, and actively encourages alternative strategies, users will
need to factor this into their database design decisions, using either nested documents or document references. While this
reportedly improves query performance, it may add further complexity to the database design for relational database
users.

Apart from the implementation of the overall design of the IRT MongoDB database, MongoDB also provides functions for
advanced queries and calculations to be performed on database data. One example of this would be the use of the map
reduce programming model for calculations on MongoDB data. Shown below is an example of MongoDB's map reduce for
calculating the total and average distances travelled by trains on each different railway track. This requires more
advanced used of host programming language features.

\begin{lstlisting}[caption=A basic map reduce implementation to get total and average distance travelled by trains on each track.]
// Select the Track and km fields from all documents
var mapFunc = function() {
  emit(this.Track, this.km);
};

// Reduces the km values from all mapped documents
var reduceFunc = function(keyTrack, kms) {
  return Array.sum(kms);
};

// Grabs the average of the reduced km values
var finalizeFunc = function(key, reducedVal) {
  reducedVal.avg = reducedVal.qty / reducedVal.count;

  return reducedVal;
};

// Execute the map reduce with the specified functions
db.train.mapReduce(
  mapFunc,
  reduceFunc,
  {
    out: 'total_distance',      // Stores output into new 'total_distance' collection
    query: { km: { $gt: 0 }},   // Excludes bad (< 0) km values
    finalize: finalizeFunc
  }
);
\end{lstlisting}

@@

In conclusion, the Monash University Institute of Railway Technology's railway data project has been shown to be possibly
managed on a non-relational database, MongoDB. We have highlighted the differences in overall database design that have
to be taken into consideration when implementing the dataset on a non-relational database, and ways the database
handles special cases that have been proving to be a problem on their current relational model.

Overall, the implementation of their dataset on a non-relational database has been positive, without any major issues that
would impact the database design in a negative way. The one aspect that would need to be looked at further, would be a
quantitative evaluation of given queries that are commonly performed on their current relational database, compared with
the equivalent queries in MongoDB. We have shown that it is very possible and beneficial to use MongoDB with the IRT
project data, however performance is an important aspect and should be looked into further to allow an informed decision
to be made.
